### 2020-11-17 tex/LitReview/LitReview.tex
initial commit from desktop.

\documentclass[a4paper]{article}

\input{includes/preamble.tex}

\begin{document}
\title{Increasing Productivity in the Domiciliary Care Sector Through Automated Routing and Scheduling of Carers}
%\author{Asyl L. Hawa}
\date{}
\maketitle

\noindent \citet{yuan2017}: Disruption Management for the Real-Time Home Caregiver Scheduling and Routing Problem

\section{Differences}
\begin{itemize}[leftmargin=*, itemsep=-0.1em]
	\item Paper covers updating an existing schedule when an disruption occurs. The disruption in this case is when a new service is added to the schedule. However, in our case there could be a variety of reasons why a disruption may occur, such as a delay in a service or carer illness.
	\item This paper does \emph{not} consider double services or synchronised (dependent) services.
	\item The paper assumes that all carers start and end their routes at the Home Health Care Centre, rather than their own homes.
	\item The paper also states that each patient is visited by a carer exactly once, which is not always the case in our problem, as a patient may require multiple visits in a single day.
	\item 	
\end{itemize}

\bibliographystyle{plainnat}
\bibliography{includes/bibliography}

\end{document}