### 2020-11-17 tex/SocialCare/SocialCare.tex

\documentclass[a4paper]{article}

\input{includes/preamble.tex}

\begin{document}
\title{Increasing Productivity in the Domiciliary Care Sector Through Automated Routing and Scheduling of Carers}
%\author{Asyl L. Hawa}
\date{}
\maketitle

\section{Constraints}
\begin{itemize}[leftmargin=*, itemsep=-0.1em]
	\item Maximum number of carers per patient
	\item PEG feed
	\item Determine the final objective, or allow for multiple objective (e.g. reduce costs, reduce travel time, reduce tardiness etc.)
	\item Maximum distance from depot (home), 6.7 miles?
	\item Maximum number of dependent jobs/precendence relations
	\item Determine the hard and soft constraints - current assuming that all patients must be visited by qualified carers, and tardiness is allowed as a result
	\item Total number of patients and total number of carers in each location/area
	\item How often are new patients added to the system/schedule?
	\item Carers updating patients files after each service, has this time been included in the service time?
	\item Number of patients in each location that require a carer that speaks the same language and is qualified for the job
	\item PPE removal and application between services
	\item Can patients refuse `treatment' if carer is late? (non-urgent service, e.g. running errands or going outdoors)
	\item Preferences, e.g. allergies, male or female carer/patient, languages, personal preferences (some carers may not wish to work with one another on double services) 
	\item Noted that carers are assigned different patients over longer periods of time to maintain professional relationships between carers and patients
	\item Maximum travel distance for individual carer's routes?
	\item Length of day for carers (half day/full days set or can schedule be flexible)?
	\item Lunch breaks?
	\item Shadowing carers
	\item Parking availability at each patient's household, or does walking time also need to be included?
	\item Are there distinct job types?
\end{itemize}

\section{Rerouting and Rescheduling}
\begin{itemize}[leftmargin=*, itemsep=-0.em]
	\item In what scenarios does and r\&r need to occur?
	\item Addition of new patient (job) to the schedule (for example, emergency revisit)
	\item Cancellation of job
	\item Scenario where a patient that originally only required one carer now requires multiple carers
	\item Change in skill requirement of a patient, where the originally scheduled carer does not have the required skill level
	\item Patient emergency that requires the carer to remain with the patient; thus remaining jobs in the carer's route will need to be serviced by one or more other carers
	\item Carer unwell at the beginning of the day, so all jobs assigned to that carer will need to be reassigned to other carer's working that day	
\end{itemize}

\end{document}