### 2020-12-01 tex/Pres/main.tex
presentation files.

\documentclass[usenames,dvipsnames]{beamer}
\input{includes/preamble.tex}

\begin{document}

\title{Increasing Productivity in the Domiciliary Care Sector Through Automated Routing and Scheduling of Carers}
\date{}
\author{A. L. Hawa, C. Lamas-Fernandez, A. Martinez-Sykora, C. Currie, T. Monks, T. Chalklen}
\titlegraphic{\includegraphics[height=1.2cm]{figures/sotonabicare2.png}}
\maketitle

%\section{Introduction}

\begin{frame}{Task}
To create an efficient route and schedule for each carer such that all patients (jobs) are visited at the correct times by suitable carers.
\small
\begin{multicols}{2}
	\begin{itemize}
		\item Assign jobs to sufficiently skilled carer(s)
		\item Assign jobs to preferred carers(s)
		\item Assign carers to preferred jobs
		\item Workload balance %($\overline{B}$)
		\item Minimise waiting times %($\overline{W}$)
		\item Minimise tardiness %($\overline{Z}$)
		\item Minimise overtime %($\overline{O}$)
		\item Minimise travel time %($\overline{T}$)
		\item Minimise travel distance*
		%\item Assign jobs and carers according to preferences ($\overline{P}$)
	\end{itemize}	
\end{multicols}	
\end{frame}

\begin{frame}{Current Model and Approach}
%\begin{multicols}{2}
	\begin{itemize}
		\item Weighted objective function 
		\item Produce multiple schedules depending on the desired objective
		\item Currently implemented in C++ and Python
	\end{itemize}	
%\end{multicols}	
\end{frame}

\begin{frame}{Data}
This is required from each location:
\footnotesize
\begin{multicols}{2}
	\begin{itemize}
		\item Total number of patients/carers in each location
		\item Max travel distance (route/radius)
		\item Skills of carers
		\item Time window of each job
		\item Patient preferences: language, gender of carers, personal
		\item Carer preferences: gender of patient, allergies, personal
		\item Length of day for carers (full/half/flexible)
		\item Number of DS/synchronised services
		\item Distinct job types?
		\item Carers ``on call''
		\item Strength of preferences
	\end{itemize}	
\end{multicols}	
\end{frame}

\begin{frame}{Queries}
\scriptsize
%\begin{multicols}{2}
	\begin{itemize}
		\item \note{Desired objective(s)}
		\item Are there skill ``levels'' for carers?
		\item Distinct job types or customised per job
		\item Jobs that \emph{cannot} start late
		\item Lunch breaks
		\item Parking availability at patient's home
		\item PPE/COVID-19 conditions
		\item Frequency of schedule formation
		\item Database/software currently used
		\item Hard and soft constraints
		\item Any additional constraints?
		\item Who will be using the application?
		\item Additional carers available?
	\end{itemize}	
%\end{multicols}
\end{frame}

\begin{frame}{Rerouting and Rescheduling (R\&R)}
In the event of an unexpected disruption, the original plan may need to be updated to accommodate any changes that occur.

\note{Which of these do we need to address?}
\small
\begin{itemize}
	\item Cancellation of a service
	\item Addition of a new service (single or double)
	\item Modification of a service (single or double)
	\item Patient emergency, carer remains with patient
	\item Increase in skill level of service
	\item Increase in number of carers required for a service
	\item Carer unwell at beginning/during the day or personal emergency
	\item Carer requiring urgent assistance from another carer
\end{itemize}	
\end{frame}

\begin{frame}{Meeting}
\small
\begin{itemize}
	\item EasyLog only has details of time/length of job
	\item Preferences/carer skills/client needs are known by staff only
	\item No jobs between 2-4pm
	\item Day in two shifts: 7am-2pm, 4pm-10pm
	\item ``Ideal rotas'' on EasyLog
	\item Availability of extra carers (e.g. part-time, can help with issues)
	\item TW 15mins before and after start time
	\item QR code carer clock in/out, not accurate
	\item Larger gaps between jobs for new carers, e.g. +10mins travel time
	\item Continuity of care
\end{itemize}

\end{frame}

\begin{frame}{Literature}

\end{frame}

\end{document}