### 2021-01-07 tex/Pres/main.tex
WIP commit.

\documentclass[usenames,dvipsnames]{beamer}
\input{includes/preamble.tex}
\setbeamertemplate{caption}{\raggedright\insertcaption\par}

\begin{document}

\title{Increasing Productivity in the Domiciliary Care Sector Through Automated Routing and Scheduling of Carers}
\date{}
\author{A. L. Hawa, C. Lamas-Fernandez, A. Martinez-Sykora, C. Currie, T. Monks, T. Chalklen}
\titlegraphic{\includegraphics[height=1.2cm]{figures/sotonabicare2.png}}
\maketitle

%\section{Introduction}

\begin{frame}{Task}
To create an efficient route and schedule for each carer such that all patients (jobs) are visited at the correct times by suitable carers.
\small
\begin{multicols}{2}
	\begin{itemize}
		\item Assign jobs to sufficiently skilled carer(s)
		\item Assign jobs to preferred carers(s)
		\item Assign carers to preferred jobs
		\item Workload balance %($\overline{B}$)
		\item Minimise waiting times %($\overline{W}$)
		\item Minimise tardiness %($\overline{Z}$)
		\item Minimise overtime %($\overline{O}$)
		\item Minimise travel time %($\overline{T}$)
		\item Minimise travel distance*
		%\item Assign jobs and carers according to preferences ($\overline{P}$)
	\end{itemize}	
\end{multicols}	
\end{frame}

\begin{frame}{Current Model and Approach}
%\begin{multicols}{2}
	\begin{itemize}
		\item Weighted objective function 
		\item Produce multiple schedules depending on the desired objective
		\item Currently implemented in C++ and Python
	\end{itemize}	
%\end{multicols}	
\end{frame}

\begin{frame}{Data}
This is required from each location:
\footnotesize
\begin{multicols}{2}
	\begin{itemize}
		\item Total number of patients/carers in each location
		\item Max travel distance (route/radius)
		\item Skills of carers
		\item Time window of each job
		\item Patient preferences: language, gender of carers, personal
		\item Carer preferences: gender of patient, allergies, personal
		\item Length of day for carers (full/half/flexible)
		\item Number of DS/synchronised services
		\item Distinct job types?
		\item Carers ``on call''
		\item Strength of preferences
	\end{itemize}	
\end{multicols}	
\end{frame}

\begin{frame}{Queries}
\scriptsize
%\begin{multicols}{2}
	\begin{itemize}
		\item \note{Desired objective(s)}
		\item Are there skill ``levels'' for carers?
		\item Distinct job types or customised per job
		\item Jobs that \emph{cannot} start late
		\item Lunch breaks
		\item Parking availability at patient's home
		\item PPE/COVID-19 conditions
		\item Frequency of schedule formation
		\item Database/software currently used
		\item Hard and soft constraints
		\item Any additional constraints?
		\item Who will be using the application?
		\item Additional carers available?
	\end{itemize}	
%\end{multicols}
\end{frame}

\begin{frame}{Rerouting and Rescheduling (R\&R)}
In the event of an unexpected disruption, the original plan may need to be updated to accommodate any changes that occur.

\note{Which of these do we need to address?}
\small
\begin{itemize}
	\item Cancellation of a service
	\item Addition of a new service (single or double)
	\item Modification of a service (single or double)
	\item Patient emergency, carer remains with patient
	\item Increase in skill level of service
	\item Increase in number of carers required for a service
	\item Carer unwell at beginning/during the day or personal emergency
	\item Carer requiring urgent assistance from another carer
\end{itemize}	
\end{frame}

\begin{frame}{Meeting}
\small
\begin{itemize}
	\item EasyLog only has details of time/length of job
	\item Preferences/carer skills/client needs are known by staff only
	\item No jobs between 2-4pm
	\item Day in two shifts: 7am-2pm, 4pm-10pm
	\item ``Ideal rotas'' on EasyLog
	\item Availability of extra carers (e.g. part-time, can help with issues)
	\item TW 15mins before and after start time
	\item QR code carer clock in/out, not accurate
	\item Larger gaps between jobs for new carers, e.g. +10mins travel time
	\item Continuity of care
\end{itemize}

\end{frame}

\begin{frame}{Literature}

\end{frame}

\begin{frame}{Comparison}
\small
\begin{multicols}{3}	
\begin{block}{Hampshire}
	\begin{itemize}
		\item Carers = 15
		\item Clients = 34
		\item Jobs = 480
	\end{itemize}
\end{block}
\begin{block}{Monmouth}
	\begin{itemize}
		\item Carers = 23
		\item Clients = 47
		\item Jobs = 735
	\end{itemize}
\end{block}
\begin{block}{Aldershot}
	\begin{itemize}
		\item Carers = 9
		\item Clients = 19
		\item Jobs = 184
	\end{itemize}
\end{block}
\end{multicols}	
\end{frame}

\begin{frame}{Comparison: Hampshire Time and Distance}
	\begin{figure}
		\includegraphics[width=0.5\linewidth]{figures/Hampshire_time_comparison_abi}%
		\includegraphics[width=0.5\linewidth]{figures/Hampshire_dist_comparison_abi}
	\end{figure}
\end{frame}

\begin{frame}{Comparison: Monmouth Time and Distance}
	\begin{figure}
		\includegraphics[width=0.5\linewidth]{figures/Monmouth_time_comparison_abi}%
		\includegraphics[width=0.5\linewidth]{figures/Monmouth_dist_comparison_abi}
	\end{figure}
\end{frame}

\begin{frame}{Comparison: Aldershot Time and Distance}
	\begin{figure}
		\includegraphics[width=0.5\linewidth]{figures/Aldershot_time_comparison_abi}%
		\includegraphics[width=0.5\linewidth]{figures/Aldershot_dist_comparison_abi}
	\end{figure}
\end{frame}

\begin{frame}{Comparison: Hampshire Peak/Off-Peak Times}
	\begin{figure}
		\includegraphics[width=0.5\linewidth]{figures/Hampshire_peaktime_comparison_abi}%
		\includegraphics[width=0.5\linewidth]{figures/Hampshire_offpeaktime_comparison_abi}
	\end{figure}
\end{frame}

\begin{frame}{Comparison: Monmouth Peak/Off-Peak Times}
	\begin{figure}
		\includegraphics[width=0.5\linewidth]{figures/Monmouth_peaktime_comparison_abi}%
		\includegraphics[width=0.5\linewidth]{figures/Monmouth_offpeaktime_comparison_abi}
	\end{figure}
\end{frame}

\begin{frame}{Comparison: Aldershot Peak/Off-Peak Times}
	\begin{figure}
		\includegraphics[width=0.5\linewidth]{figures/Aldershot_peaktime_comparison_abi}%
		\includegraphics[width=0.5\linewidth]{figures/Aldershot_offpeaktime_comparison_abi}
	\end{figure}
\end{frame}

\begin{frame}
	\begin{figure}
		\includegraphics[width=0.5\textwidth]{figures/Hampshire_timeofday_jobs_abi}
	\end{figure}
	\vspace*{-0.9cm}
	\begin{columns}[T]
		\begin{column}{0.5\textwidth}
			\begin{figure}
				\includegraphics[width=\textwidth]{figures/Monmouth_timeofday_jobs_abi}
			\end{figure}
		\end{column}%
		\begin{column}{0.5\textwidth}
			\begin{figure}
				\includegraphics[width=\textwidth]{figures/Aldershot_timeofday_jobs_abi}
			\end{figure}
		\end{column}
	\end{columns}
\end{frame}

\begin{frame}
	\begin{figure}
		\includegraphics[width=0.5\textwidth]{figures/Hampshire_shift_jobs_abi}
	\end{figure}
	\vspace*{-0.8cm}
	\begin{columns}[T]
		\begin{column}{0.5\textwidth}
			\begin{figure}
				\includegraphics[width=\textwidth]{figures/Monmouth_shift_jobs_abi}
			\end{figure}
		\end{column}%
		\begin{column}{0.5\textwidth}
			\begin{figure}
				\includegraphics[width=\textwidth]{figures/Aldershot_shift_jobs_abi}
			\end{figure}
		\end{column}
	\end{columns}
\end{frame}

\begin{frame}{Comparison: Total Time (minutes)}
\scriptsize
\begin{table}
%\caption{Comparison}
\renewcommand{\arraystretch}{1.5}
\begin{tabular}{ccrrcrrcrr}\toprule
	&& \multicolumn{2}{c}{Hampshire} && \multicolumn{2}{c}{Monmouth} && \multicolumn{2}{c}{Aldershot} \\
	\cmidrule{3-4} \cmidrule{6-7} \cmidrule{9-10}
	\multicolumn{1}{c}{Date} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} \\
	\midrule
	02-Nov && 5640.00 & \gre{5452.46} && 6285.00 & \gre{5851.16} && 1480.00 & \gre{1449.90}\\
	03-Nov && 5460.00 & \gre{5282.68} && 6030.00 & \gre{5758.70} && 1080.00 & \gre{1015.74}\\
	04-Nov && 5160.00 & \gre{4996.38} && 6135.00 & \gre{6044.82} && 1355.00 & \gre{1321.32}\\
	05-Nov && 5265.00 & \gre{5089.72} && 7200.00 & \gre{6955.46} && 2000.00 & \gre{1905.00}\\
	06-Nov && 4875.00 & \gre{4617.36} && 7455.00 & \gre{7100.96} && 1470.00 & \gre{1345.28}\\
	07-Nov && 3780.00 & \gre{3616.35} && 4845.00 & \gre{4574.79} && 1065.00 & \gre{1005.92}\\
	08-Nov && 3720.00 & \gre{3608.01} && 4470.00 & \gre{4380.78} && 1050.00 & \gre{1006.87}\\
	\bottomrule
\end{tabular}
\end{table}%
\end{frame}

\begin{frame}{Comparison: Total Travel Time (minutes)}
\scriptsize
\begin{table}
%\caption{Comparison}
\renewcommand{\arraystretch}{1.5}
\begin{tabular}{ccrrcrrcrr}\toprule
	&& \multicolumn{2}{c}{Hampshire} && \multicolumn{2}{c}{Monmouth} && \multicolumn{2}{c}{Aldershot} \\
	\cmidrule{3-4} \cmidrule{6-7} \cmidrule{9-10}
	\multicolumn{1}{c}{Date} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} \\
	\midrule
	02-Nov && 475.00 & \gre{232.28} && 722.55 & \gre{687.00} && 192.77 & \gre{124.81}\\
	03-Nov && 489.88 & \gre{227.02} && \red{728.62} & 745.88 && 166.05 & \gre{119.21}\\
	04-Nov && 490.77 & \gre{242.58} && \red{769.78} & 895.18 && 178.28 & \gre{119.79}\\
	05-Nov && 490.42 & \gre{271.44} && 798.20 & \gre{687.87} && 211.58 & \gre{122.75}\\
	06-Nov && 479.50 & \gre{243.76} && \red{864.52} & 920.79 && 152.97 & \gre{108.11}\\
	07-Nov && 452.67 & \gre{243.81} && \red{694.07} & 779.27 && 167.70 & \gre{108.08}\\
	08-Nov && 405.41 & \gre{223.68} && \red{666.17} & 802.96 && 176.48 & \gre{119.98}\\
	\bottomrule
\end{tabular}
\end{table}%
\end{frame}

\begin{frame}{Comparison NEW: Total Travel Time (minutes)}
	\scriptsize
	\begin{table}
		%\caption{Comparison}
		\renewcommand{\arraystretch}{1.5}
		\begin{tabular}{ccrrcrrcrr}\toprule
			&& \multicolumn{2}{c}{Hampshire} && \multicolumn{2}{c}{Monmouth} && \multicolumn{2}{c}{Aldershot} \\
			\cmidrule{3-4} \cmidrule{6-7} \cmidrule{9-10}
			\multicolumn{1}{c}{Date} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} \\
			\midrule
			02-Nov && 372.88 & \gre{232.28} && \red{674.46} & 687.00 && 153.05 & \gre{124.81}\\
			03-Nov && 381.58 & \gre{227.02} && \red{675.89} & 745.88 && 129.47 & \gre{119.21}\\
			04-Nov && 384.52 & \gre{242.58} && \red{722.42} & 895.18 && 141.19 & \gre{119.79}\\
			05-Nov && 389.95 & \gre{271.44} && 745.86 & \gre{687.87} && 168.99 & \gre{122.7}5\\
			06-Nov && 378.22 & \gre{243.76} && \red{824.72} & 920.79 && 123.08 & \gre{108.11}\\
			07-Nov && 356.83 & \gre{243.81} && \red{661.50} & 779.27 && 133.91 & \gre{108.08}\\
			08-Nov && 324.72 & \gre{223.68} && \red{636.07} & 802.96 && 142.60 & \gre{119.98}\\
			\bottomrule
		\end{tabular}
	\end{table}%
\end{frame}

\begin{frame}{Comparison: Total Waiting Time (minutes)}
\scriptsize
\begin{table}
%\caption{Comparison}
\renewcommand{\arraystretch}{1.5}
\begin{tabular}{ccrrcrrcrr}\toprule
	&& \multicolumn{2}{c}{Hampshire} && \multicolumn{2}{c}{Monmouth} && \multicolumn{2}{c}{Aldershot} \\
	\cmidrule{3-4} \cmidrule{6-7} \cmidrule{9-10}
	\multicolumn{1}{c}{Date} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} \\
	\midrule
	02-Nov && \red{1595.00} & 1650.18 && 1407.45 & \gre{1009.16} && \red{222.23} & 260.09\\
	03-Nov && \red{1580.12} & 1665.65 && 1026.38 & \gre{737.81} && 73.95 & \gre{56.53}\\
	04-Nov && \red{1189.23} & 1273.80 && 775.22 & \gre{559.64} && \red{36.72} & 61.53\\
	05-Nov && \red{1159.58} & 1203.28 && 1661.80 & \gre{1527.59} && 338.42 & \gre{332.25}\\
	06-Nov && 1065.50 & \gre{1043.60} && 1505.48 & \gre{1095.16} && 402.03 & \gre{322.17}\\
	07-Nov && \red{957.33} & 1002.54 && 790.93 & \gre{435.51} && \red{212.30} & 212.84\\
	08-Nov && \red{1019.59} & 1089.32 && 548.83 & \gre{322.82} && \red{143.52} & 156.89\\
	\bottomrule
\end{tabular}
\end{table}%
\end{frame}

\begin{frame}{Comparison NEW: Total Waiting Time (minutes)}
	\scriptsize
	\begin{table}
		%\caption{Comparison}
		\renewcommand{\arraystretch}{1.5}
		\begin{tabular}{ccrrcrrcrr}\toprule
			&& \multicolumn{2}{c}{Hampshire} && \multicolumn{2}{c}{Monmouth} && \multicolumn{2}{c}{Aldershot} \\
			\cmidrule{3-4} \cmidrule{6-7} \cmidrule{9-10}
			\multicolumn{1}{c}{Date} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} \\
			\midrule
			02-Nov && 1697.12 & \gre{1650.18} && 1455.54 & \gre{1009.16} && 261.95 & \gre{260.09}\\
			03-Nov && 1688.42 & \gre{1665.65} && 1079.11 & \gre{737.81} && 110.53 & \gre{56.53}\\
			04-Nov && 1295.48 & \gre{1273.80} && 822.58 & \gre{559.64} && 73.81 & \gre{61.53}\\
			05-Nov && 1260.05 & \gre{1203.28} && 1714.14 & \gre{1527.59} && 381.01 & \gre{332.25}\\
			06-Nov && 1166.78 & \gre{1043.60} && 1545.28 & \gre{1095.16} && 431.92 & \gre{322.17}\\
			07-Nov && 1053.17 & \gre{1002.54} && 823.50 & \gre{435.51} && 246.09 & \gre{212.84}\\
			08-Nov && 1100.28 & \gre{1089.32} && 578.93 & \gre{322.82} && 177.41 & \gre{156.89}\\
			\bottomrule
		\end{tabular}
	\end{table}%
\end{frame}

\begin{frame}{Comparison: Total Travel Distance (km)}
	\scriptsize
	\begin{table}
		%\caption{Comparison}
		\renewcommand{\arraystretch}{1.5}
		\begin{tabular}{ccrrcrrcrr}\toprule
			&& \multicolumn{2}{c}{Hampshire} && \multicolumn{2}{c}{Monmouth} && \multicolumn{2}{c}{Aldershot} \\
			\cmidrule{3-4} \cmidrule{6-7} \cmidrule{9-10}
			\multicolumn{1}{c}{Date} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} \\
			\midrule
			02-Nov && 288.26 & \gre{161.16} && \red{497.03} & 569.09 && 127.14 & \gre{100.00}\\
			03-Nov && 307.63 & \gre{153.31} && \red{502.23} & 662.55 && 106.41 & \gre{99.03}\\
			04-Nov && 301.65 & \gre{165.69} && \red{550.60} & 822.70 && 108.11 & \gre{90.21}\\
			05-Nov && 302.08 & \gre{193.55} && \red{561.90} & 578.20 && 129.57 & \gre{92.00}\\
			06-Nov && 301.39 & \gre{167.34} && \red{639.99} & 831.14 && 98.11 & \gre{83.70}\\
			07-Nov && 291.58 & \gre{176.95} && \red{503.89} & 662.54 && 110.56 & \gre{82.77}\\
			08-Nov && 250.77 & \gre{160.10} && \red{489.51} & 676.26 && 113.47 & \gre{91.53}\\
			\bottomrule
		\end{tabular}
	\end{table}%
\end{frame}

\begin{comment}
\begin{frame}{Comparison}
	\begin{figure}[T]
		\includegraphics[width=0.5\linewidth]{figures/2_Nov_20_Hampshire_time_info_abi}%
		\includegraphics[width=0.5\linewidth]{figures/2_Nov_20_Hampshire_time_info_dst}
	\end{figure}
\end{frame}

\begin{frame}{Comparison}
	\begin{figure}[T]
		\includegraphics[width=0.5\linewidth]{figures/3_Nov_20_Hampshire_time_info_abi}%
		\includegraphics[width=0.5\linewidth]{figures/3_Nov_20_Hampshire_time_info_dst}
	\end{figure}
\end{frame}

\begin{frame}{Comparison}
	\begin{figure}[T]
		\includegraphics[width=0.5\linewidth]{figures/4_Nov_20_Hampshire_time_info_abi}%
		\includegraphics[width=0.5\linewidth]{figures/4_Nov_20_Hampshire_time_info_dst}
	\end{figure}
\end{frame}

\begin{frame}{Comparison}
	\begin{figure}[T]
		\includegraphics[width=0.5\linewidth]{figures/5_Nov_20_Hampshire_time_info_abi}%
		\includegraphics[width=0.5\linewidth]{figures/5_Nov_20_Hampshire_time_info_dst}
	\end{figure}
\end{frame}

\begin{frame}{Comparison}
	\begin{figure}[T]
		\includegraphics[width=0.5\linewidth]{figures/6_Nov_20_Hampshire_time_info_abi}%
		\includegraphics[width=0.5\linewidth]{figures/6_Nov_20_Hampshire_time_info_dst}
	\end{figure}
\end{frame}

\begin{frame}{Comparison}
	\begin{figure}[T]
		\includegraphics[width=0.5\linewidth]{figures/7_Nov_20_Hampshire_time_info_abi}%
		\includegraphics[width=0.5\linewidth]{figures/7_Nov_20_Hampshire_time_info_dst}
	\end{figure}
\end{frame}

\begin{frame}{Comparison}
	\begin{figure}[T]
		\includegraphics[width=0.5\linewidth]{figures/8_Nov_20_Hampshire_time_info_abi}%
		\includegraphics[width=0.5\linewidth]{figures/8_Nov_20_Hampshire_time_info_dst}
	\end{figure}
\end{frame}

\begin{frame}{Comparison}
	\begin{figure}
		\includegraphics[width=0.5\linewidth]{figures/2_Nov_20_Hampshire_workload_abi}%
		\includegraphics[width=0.5\linewidth]{figures/2_Nov_20_Hampshire_workload_dst}
	\end{figure}
\end{frame}

\begin{frame}{Comparison}
	\begin{figure}
		\includegraphics[width=0.5\linewidth]{figures/3_Nov_20_Hampshire_workload_abi}%
		\includegraphics[width=0.5\linewidth]{figures/3_Nov_20_Hampshire_workload_dst}
	\end{figure}
\end{frame}

\begin{frame}{Comparison}
	\begin{figure}
		\includegraphics[width=0.5\linewidth]{figures/4_Nov_20_Hampshire_workload_abi}%
		\includegraphics[width=0.5\linewidth]{figures/4_Nov_20_Hampshire_workload_dst}
	\end{figure}
\end{frame}

\begin{frame}{Comparison}
	\begin{figure}
		\includegraphics[width=0.5\linewidth]{figures/5_Nov_20_Hampshire_workload_abi}%
		\includegraphics[width=0.5\linewidth]{figures/5_Nov_20_Hampshire_workload_dst}
	\end{figure}
\end{frame}

\begin{frame}{Comparison}
	\begin{figure}
		\includegraphics[width=0.5\linewidth]{figures/6_Nov_20_Hampshire_workload_abi}%
		\includegraphics[width=0.5\linewidth]{figures/6_Nov_20_Hampshire_workload_dst}
	\end{figure}
\end{frame}

\begin{frame}{Comparison}
	\begin{figure}
		\includegraphics[width=0.5\linewidth]{figures/7_Nov_20_Hampshire_workload_abi}%
		\includegraphics[width=0.5\linewidth]{figures/7_Nov_20_Hampshire_workload_dst}
	\end{figure}
\end{frame}

\begin{frame}{Comparison}
	\begin{figure}
		\includegraphics[width=0.5\linewidth]{figures/8_Nov_20_Hampshire_workload_abi}%
		\includegraphics[width=0.5\linewidth]{figures/8_Nov_20_Hampshire_workload_dst}
	\end{figure}
\end{frame}
\end{comment}

\begin{comment}
%SAVE TEMPLATE
\begin{frame}{Comparison: Total Time (minutes)}
\scriptsize
\begin{table}
%\caption{Comparison}
\renewcommand{\arraystretch}{1.5}
\begin{tabular}{ccrrcrrcrr}\toprule
&& \multicolumn{2}{c}{Hampshire} && \multicolumn{2}{c}{Monmouth} && \multicolumn{2}{c}{Aldershot} \\
\cmidrule{3-4} \cmidrule{6-7} \cmidrule{9-10}
\multicolumn{1}{c}{Date} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} \\
\midrule
02-Nov && 5640.00 & 5452.46 && 6285.00 & 5851.16 && 1480.00 & 1449.90\\
03-Nov && 5460.00 & 5282.68 && 6030.00 & 5758.70 && 1080.00 & 1015.74\\
04-Nov && 5160.00 & 4996.38 && 6135.00 & 6044.82 && 1355.00 & 1321.32\\
05-Nov && 5265.00 & 5089.72 && 7200.00 & 6955.46 && 2000.00 & 1905.00\\
06-Nov && 4875.00 & 4617.36 && 7455.00 & 7100.96 && 1470.00 & 1345.28\\
07-Nov && 3780.00 & 3616.35 && 4845.00 & 4574.79 && 1065.00 & 1005.92\\
08-Nov && 3720.00 & 3608.01 && 4470.00 & 4380.78 && 1050.00 & 1006.87\\
\bottomrule
\end{tabular}
\end{table}%
\end{frame}

\begin{frame}{Comparison: Total Travel Time (minutes)}
\scriptsize
\begin{table}
%\caption{Comparison}
\renewcommand{\arraystretch}{1.5}
\begin{tabular}{ccrrcrrcrr}\toprule
&& \multicolumn{2}{c}{Hampshire} && \multicolumn{2}{c}{Monmouth} && \multicolumn{2}{c}{Aldershot} \\
\cmidrule{3-4} \cmidrule{6-7} \cmidrule{9-10}
\multicolumn{1}{c}{Date} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} \\
\midrule
02-Nov && 475.00 & 232.28 && 722.55 & 687.00 && 192.77 & 124.81\\
03-Nov && 489.88 & 227.02 && 728.62 & 745.88 && 166.05 & 119.21\\
04-Nov && 490.77 & 242.58 && 769.78 & 895.18 && 178.28 & 119.79\\
05-Nov && 490.42 & 271.44 && 798.20 & 687.87 && 211.58 & 122.75\\
06-Nov && 479.50 & 243.76 && 864.52 & 920.79 && 152.97 & 108.11\\
07-Nov && 452.67 & 243.81 && 694.07 & 779.27 && 167.70 & 108.08\\
08-Nov && 405.41 & 223.68 && 666.17 & 802.96 && 176.48 & 119.98\\
\bottomrule
\end{tabular}
\end{table}%
\end{frame}

\begin{frame}{Comparison: Total Waiting Time (minutes)}
	\scriptsize
	\begin{table}
		%\caption{Comparison}
		\renewcommand{\arraystretch}{1.5}
		\begin{tabular}{ccrrcrrcrr}\toprule
			&& \multicolumn{2}{c}{Hampshire} && \multicolumn{2}{c}{Monmouth} && \multicolumn{2}{c}{Aldershot} \\
			\cmidrule{3-4} \cmidrule{6-7} \cmidrule{9-10}
			\multicolumn{1}{c}{Date} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} \\
			\midrule
			02-Nov && 1595.00 & 1650.18 && 1407.45 & 1009.16 && 222.23 & 260.09\\
			03-Nov && 1580.12 & 1665.65 && 1026.38 & 737.81 && 73.95 & 56.53\\
			04-Nov && 1189.23 & 1273.80 && 775.22 & 559.64 && 36.72 & 61.53\\
			05-Nov && 1159.58 & 1203.28 && 1661.80 & 1527.59 && 338.42 & 332.25\\
			06-Nov && 1065.50 & 1043.60 && 1505.48 & 1095.16 && 402.03 & 322.17\\
			07-Nov && 957.33 & 1002.54 && 790.93 & 435.51 && 212.30 & 212.84\\
			08-Nov && 1019.59 & 1089.32 && 548.83 & 322.82 && 143.52 & 156.89\\
			\bottomrule
		\end{tabular}
	\end{table}%
\end{frame}
\end{comment}

\begin{comment}
\begin{frame}
\begin{figure}
\includegraphics[width=0.5\textwidth]{figures/Hampshire_timeofday_jobs_abi}
\end{figure}
\vspace*{-0.9cm}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{figure}
\includegraphics[width=\textwidth]{figures/Monmouth_timeofday_jobs_abi}
\end{figure}
\end{column}%
\begin{column}{0.5\textwidth}
\begin{figure}
\includegraphics[width=\textwidth]{figures/Aldershot_timeofday_jobs_abi}
\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}
\begin{figure}
\includegraphics[width=0.5\textwidth]{figures/Hampshire_shift_jobs_abi}
\end{figure}
\vspace*{-0.8cm}
\begin{columns}[T]
\begin{column}{0.5\textwidth}
\begin{figure}
\includegraphics[width=\textwidth]{figures/Monmouth_shift_jobs_abi}
\end{figure}
\end{column}%
\begin{column}{0.5\textwidth}
\begin{figure}
\includegraphics[width=\textwidth]{figures/Aldershot_shift_jobs_abi}
\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}{Comparison}
\begin{figure}[T]
\includegraphics[width=0.5\linewidth]{figures/2_Nov_20_Hampshire_time_info_abi}%
\includegraphics[width=0.5\linewidth]{figures/2_Nov_20_Hampshire_time_info_dst}
\end{figure}
\end{frame}

\begin{frame}{Comparison}
\begin{figure}
\includegraphics[width=0.5\linewidth]{figures/2_Nov_20_Hampshire_workload_abi}%
\includegraphics[width=0.5\linewidth]{figures/2_Nov_20_Hampshire_workload_dst}
\end{figure}
\end{frame}

\begin{frame}{Comparisonx}
\begin{figure}
\includegraphics[width=0.5\linewidth]{figures/3_Nov_20_Hampshire_workload_abi}%
\includegraphics[width=0.5\linewidth]{figures/3_Nov_20_Hampshire_workload_dst}
\end{figure}
\end{frame}

\begin{frame}{Comparison}
\begin{figure}[B]
\includegraphics[width=0.5\linewidth]{figures/2_Nov_20_Monmouth_time_info_abi}%
\includegraphics[width=0.5\linewidth]{figures/2_Nov_20_Monmouth_workload_abi}
\end{figure}
\end{frame}

\begin{frame}{Comparison}
\begin{figure}
\includegraphics[width=0.5\linewidth]{figures/2_Nov_20_Aldershot_time_info_abi}%
\includegraphics[width=0.5\linewidth]{figures/2_Nov_20_Aldershot_workload_abi}
\end{figure}
\end{frame}

\begin{frame}
\begin{columns}
\column{0.5\textwidth}
\begin{minipage}[c][0.4\textheight][c]{\linewidth}
\centering
\includegraphics[width=0.9\linewidth]{figures/Hampshire_timeofday_jobs_abi}
\end{minipage}
\begin{minipage}[c][0.4\textheight][c]{\linewidth}
\centering
\includegraphics[width=0.9\linewidth]{figures/Monmouth_timeofday_jobs_abi}
\end{minipage}
\column{0.5\textwidth}
\begin{minipage}[c][0.4\textheight][c]{\linewidth}
\begin{enumerate}
\item First.
\item Second.
\item Third.
\end{enumerate}
\end{minipage}
\begin{minipage}[c][0.4\textheight][c]{\linewidth}
\centering
\includegraphics[width=0.9\linewidth]{figures/Aldershot_timeofday_jobs_abi}
\end{minipage}
\end{columns}
\end{frame}

\begin{frame}
\scriptsize
\begin{multicols}{3}
\begin{table}
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.5}
%		\caption{Hampshire}
\begin{tabular}{ccc}\toprule
Date & Abicare & DST \\
\midrule
02-Nov & 5640.00 & 5452.46\\
03-Nov & 5460.00 & 5282.68\\
04-Nov & 5160.00 & 4996.38\\
05-Nov & 5265.00 & 5089.72\\
06-Nov & 4875.00 & 4617.36 \\
07-Nov & 3780.00 & 3616.35\\
08-Nov & 3720.00 & 3608.01\\
\bottomrule
\end{tabular}
\end{table}%
\vspace{2mm}
\begin{table}
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.5}
\begin{tabular}{ccc}\toprule
Date & Abicare & DST \\
\midrule
02-Nov & 6285.00 & 5851.16\\
03-Nov & 6030.00 & 5758.70\\
04-Nov & 6135.00 & 6044.82\\
05-Nov & 7200.00 & 6955.46\\
06-Nov & 7455.00 & 7100.96 \\
07-Nov & 4845.00 & 4574.79\\
08-Nov & 4470.00 & 4380.78\\
\bottomrule
\end{tabular}
\end{table}%
\vspace{2mm}
\begin{table}
\setlength{\tabcolsep}{3pt}
\renewcommand{\arraystretch}{1.5}
\begin{tabular}{ccc}\toprule
Date & Abicare & DST \\
\midrule
02-Nov & 1480.00 & 1449.90\\
03-Nov & 1080.00 & 1015.74\\
04-Nov & 1355.00 & 1321.32\\
05-Nov & 2000.00 & 1905.00\\
06-Nov & 1470.00 & 1345.28 \\
07-Nov & 1065.00 & 1005.92\\
08-Nov & 1050.00 & 1006.87\\
\bottomrule
\end{tabular}
\end{table}
\end{multicols}
\end{frame}
\end{comment}

\end{document}