### 2021-06-28 tex/Documentation/Documentation.tex
Documentation for python program abicare.

\documentclass[a4paper]{article}

\input{includes/preamble.tex}

\begin{document}
\title{Documentation: Abicare DST}
%\author{Asyl L. Hawa}
\date{}
\maketitle

\section{Disruption Management for the RT-HCSRP}

\subsection{Differences}
\begin{itemize}[label=\textcolor{myPurple}{\textbullet},leftmargin=*,itemsep=-0.1em]
	\item 
\end{itemize}

\subsection{Model and Solution Approach}
\begin{itemize}[label=\textcolor{myPurple}{\textbullet},leftmargin=*,itemsep=-0.1em]
	\item One
	\begin{enumerate}[label=(\roman*),itemsep=-0.1em]
		\item 
	\end{enumerate}
	\item Two
	\begin{enumerate}[label=(\roman*),itemsep=-0.1em]
		\item
	\end{enumerate}
	\item 
\end{itemize}

\subsection{Modifications}
\begin{itemize}[label=\textcolor{myPurple}{\textbullet},leftmargin=*,itemsep=-0.1em]
	\item 
\end{itemize}

\section{Disruption Management of the VRP with Breakdown}

\subsection{Differences}
\begin{itemize}[label=\textcolor{myYellow}{\textbullet},leftmargin=*,itemsep=-0.1em]
	\item 
\end{itemize}

\subsection{Model and Solution Approach}
\begin{itemize}[label=\textcolor{myYellow}{\textbullet},leftmargin=*,itemsep=-0.1em]
	\item 
\end{itemize}

\subsection{Modifications}
\begin{itemize}[label=\textcolor{myYellow}{\textbullet},leftmargin=*,itemsep=-0.1em]
	\item 
\end{itemize}

\end{document}