### 2021-07-22 tex/Paper/bib_section.tex

\begin{thebibliography}{}

\bibitem[{Ait Haddadene} et~al., 2016]{AitHaddadene2016}
{Ait Haddadene}, S.~R., Labadie, N., and Prodhon, C. (2016).
\newblock {A GRASP × ILS for the vehicle routing problem with time windows,
  synchronization and precedence constraints}.
\newblock {\em Expert Systems with Applications}, 66:274--294.

\bibitem[Bachouch et~al., 2011]{Bachouch2011}
Bachouch, R.~B., Guinet, A., and Hajri-Gabouj, S. (2011).
\newblock {A Decision-Making Tool for Home Health Care Nurses' Planning}.
\newblock {\em Supply Chain Forum: An International Journal}, 12(1):14--20.

\bibitem[Begur et~al., 1997]{Begur1997}
Begur, S.~V., Miller, D.~M., and Weaver, J.~R. (1997).
\newblock {An Integrated Spatial DSS for Scheduling and Routing
  Home-Health-Care Nurses}.
\newblock {\em Interfaces}, 27(4):35--48.

\bibitem[Bertels and Fahle, 2006]{BERTELS20062866}
Bertels, S. and Fahle, T. (2006).
\newblock {A hybrid setup for a hybrid scenario: combining heuristics for the
  home health care problem}.
\newblock {\em Computers {\&} Operations Research}, 33(10):2866--2890.

\bibitem[Bredstr{\"{o}}m and R{\"{o}}nnqvist, 2008]{Bredstrom2008}
Bredstr{\"{o}}m, D. and R{\"{o}}nnqvist, M. (2008).
\newblock {Combined vehicle routing and scheduling with temporal precedence and
  synchronization constraints}.
\newblock {\em European Journal of Operational Research}, 191(1):19--29.

\bibitem[Ciss{\'{e}} et~al., 2017]{Cisse2017}
Ciss{\'{e}}, M., Yal{\c{c}}ındağ, S., Kergosien, Y., Şahin, E., Lent{\'{e}},
  C., and Matta, A. (2017).
\newblock {OR problems related to Home Health Care: A review of relevant
  routing and scheduling problems}.

\bibitem[Clarke and Wright, 1964]{Clarke1964}
Clarke, G. and Wright, J.~W. (1964).
\newblock {Scheduling of Vehicles from a Central Depot to a Number of Delivery
  Points}.
\newblock {\em Operations Research}, 12(4):568--581.

\bibitem[Decerle et~al., 2018]{Decerle2018}
Decerle, J., Grunder, O., {Hajjam El Hassani}, A., and Barakat, O. (2018).
\newblock {A memetic algorithm for a home health care routing and scheduling
  problem}.
\newblock {\em Operations Research for Health Care}, 16:59--71.

\bibitem[Drexl, 2012]{Drexl2012}
Drexl, M. (2012).
\newblock {Synchronization in Vehicle Routing—A Survey of VRPs with Multiple
  Synchronization Constraints}.
\newblock {\em Transportation Science}, 46(3):297--316.

\bibitem[Eveborn et~al., 2006]{Eveborn2006}
Eveborn, P., Flisberg, P., and R{\"{o}}nnqvist, M. (2006).
\newblock {Laps Care—an operational system for staff planning of home care}.
\newblock {\em European Journal of Operational Research}, 171(3):962--976.

\bibitem[Eveborn et~al., 2009]{Eveborn2009a}
Eveborn, P., R{\"{o}}nnqvist, M., Einarsd{\'{o}}lttir, H., Eklund, M.,
  Lid{\'{e}}n, K., and Almroth, M. (2009).
\newblock {Operations research improves quality and efficiency in home care}.
\newblock {\em Interfaces}, 39(1):18--34.

\bibitem[Feo and Resende, 1989]{FEO198967}
Feo, T.~A. and Resende, M. G.~C. (1989).
\newblock {A probabilistic heuristic for a computationally difficult set
  covering problem}.
\newblock {\em Operations Research Letters}, 8(2):67--71.

\bibitem[Fikar and Hirsch, 2015]{Fikar2015}
Fikar, C. and Hirsch, P. (2015).
\newblock {A matheuristic for routing real-world home service transport systems
  facilitating walking}.
\newblock {\em Journal of Cleaner Production}, 105:300--310.

\bibitem[Fikar and Hirsch, 2017]{Fikar2017}
Fikar, C. and Hirsch, P. (2017).
\newblock {Home health care routing and scheduling: A review}.
\newblock {\em Computers {\&} Operations Research}, 77:86--95.

\bibitem[Hansen et~al., 2010]{hansen2010variable}
Hansen, P., Mladenovi{\'{c}}, N., and P{\'{e}}rez, J. A.~M. (2010).
\newblock {Variable neighbourhood search: methods and applications}.
\newblock {\em Annals of Operations Research}, 175(1):367--407.

\bibitem[Laguna and Marti, 1999]{laguna1999grasp}
Laguna, M. and Marti, R. (1999).
\newblock {GRASP and path relinking for 2-layer straight line crossing
  minimization}.
\newblock {\em INFORMS Journal on Computing}, 11(1):44--52.

\bibitem[L{\'{o}}pez-Ib{\'{a}}{\~{n}}ez et~al., 2016]{Lopez-Ibanez2016}
L{\'{o}}pez-Ib{\'{a}}{\~{n}}ez, M., Dubois-Lacoste, J., {P{\'{e}}rez
  C{\'{a}}ceres}, L., Birattari, M., and St{\"{u}}tzle, T. (2016).
\newblock {The irace package: Iterated racing for automatic algorithm
  configuration}.
\newblock {\em Operations Research Perspectives}, 3:43--58.

\bibitem[Mankowska et~al., 2014]{Mankowska2014}
Mankowska, D.~S., Meisel, F., and Bierwirth, C. (2014).
\newblock {The home health care routing and scheduling problem with
  interdependent services}.
\newblock {\em Health Care Management Science}, 17(1):15--30.

\bibitem[Mladenovi{\'{c}} and Hansen, 1997]{Mladenovic1997}
Mladenovi{\'{c}}, N. and Hansen, P. (1997).
\newblock {Variable neighborhood search}.
\newblock {\em Computers {\&} Operations Research}, 24(11):1097--1100.

\bibitem[Mosquera et~al., 2018]{Mosquera2018}
Mosquera, F., Smet, P., and {Vanden Berghe}, G. (2018).
\newblock {Flexible home care scheduling}.
\newblock {\em Omega}.

\bibitem[{NHS England}, 2019]{england2019nhs}
{NHS England} (2019).
\newblock {The NHS long term plan}.

\bibitem[{NHS Providers}, 2018]{NHSProviders2018}
{NHS Providers} (2018).
\newblock {Community services: taking centre stage}.

\bibitem[OECD, 2019]{OECDHealthataGlance2019}
OECD (2019).
\newblock {\em {Health at a Glance 2019}}.

\bibitem[Rasmussen et~al., 2012]{Rasmussen2012}
Rasmussen, M.~S., Justesen, T., Dohn, A., and Larsen, J. (2012).
\newblock {The Home Care Crew Scheduling Problem: Preference-based visit
  clustering and temporal dependencies}.
\newblock {\em European Journal of Operational Research}, 219(3):598--610.

\bibitem[Resende and Ribeiro, 2016]{resende2016optimization}
Resende, M. G.~C. and Ribeiro, C.~C. (2016).
\newblock {\em {Optimization by GRASP}}.
\newblock Springer.

\bibitem[Rest and Hirsch, 2016]{Rest2016}
Rest, K.-D. and Hirsch, P. (2016).
\newblock {Daily scheduling of home health care services using time-dependent
  public transport}.
\newblock {\em Flexible Services and Manufacturing Journal}, 28(3):495--525.

\bibitem[Robertson et~al., 2017]{robertson2017understanding}
Robertson, R., Wenzel, L., Thompson, J., and Charles, A. (2017).
\newblock {Understanding NHS financial pressures}.
\newblock {\em The King's Fund}.

\bibitem[Tarricone and Tsouros, 2008]{tarricone2008home}
Tarricone, R. and Tsouros, A.~D. (2008).
\newblock {\em {Home care in Europe: the solid facts}}.
\newblock WHO Regional Office Europe.

\end{thebibliography}