### 2021-10-25 tex/ARCBoard/ArcBoard.tex
paragraph for ARC Board.

\documentclass[a4paper]{article}

\input{includes/preamble.tex}

\begin{document}
\title{Improving Community Health Care Logistics Using OR -- ARC Board}
%\author{Asyl L. Hawa}
\date{}
\maketitle

\noindent Throughout this project we worked alongside Abicare, a care home company based in various regions in the south of England and Wales, in order to improve upon the routing and scheduling of nurses to patients prefering domiciliary care. Business-as-usual planning at Abicare is manual, requiring staff in each area to create schedules for nurses by hand. The schedule must ensure that patients' needs are met, whilst also minimising the time and distance travelled by the nurses to each location; additional travel incurs extra costs. Furthermore, it is important to consider synchronised jobs, where the care of a patient requires two nurses, and the workload balance of each nurse. We therefore developed a Decision Support Tool which utilises a GRASP-VNS algorithm to minimise waiting times, travel times, mileage and tardiness for each of the nurses. This DST will be used in place of manual planning at Abicare, which will allow for the production of high-quality schedules and routes. In preliminary tests, which assessed instances in Basingstoke, Monmouth, and Aldershot, the DST was able to reduce the total time spent travelling by nurses in a single day up to two hours and 34 minutes, the total waiting time by up to seven hours and 30 minutes, and the total distance travelled by up to 154km. These encouraging results show that the DST will greatly benefit Abicare and the domiciliary care industry, allowing staff to focus on providing quality care to patients.

\end{document}