### 2021-10-25 tex/Plan/Plan.tex
document for plan for remainder of ARC Board.

\documentclass[a4paper]{article}

\input{includes/preamble.tex}

\begin{document}
\title{Updated Project Plan}
%\author{Asyl L. Hawa}
\date{\today}
\maketitle

\section*{Objectives}
\begin{enumerate}[leftmargin=*]
	\item \textbf{Real Evaluation of the DST}\\
	This will be done over a period of time in two ways: (a) comparing the automated plans generated by the DST with plans created manually by staff; and (b) obtaining and assessing reviews from nurses regaring the new automated plans. This will provide us with a clearer understanding of the impact of the DST in practice.
	\item \textbf{Exact Methods for the RSP}\\
	An exact method would provide a guarantee that the problem is solved to optimality, however, it may not be suitable for daily planning due to the time and software that may be required. An exact method would be useful for Health Services for considering whether investments in planning are helpful or necessary, and will also be useful as a basis for other types of modelling.
	\item \textbf{Stochastic Demand}\\
	We will identify particular situations that require modifications to existing plans, e.g. changes in the number of nurses or new patients requiring treatment, and update the model to take these unplanned changes into consideration.
\end{enumerate}

%\begin{itemize}
%	\item Compare results from original plan with new DST plans at Abicare
%	\item start looking at exact methods
%	\item look at stochastic method
%	\item publish code and paper
%\end{itemize}

\section*{Outputs}
\begin{enumerate}[leftmargin=*]
	\item \textbf{Decision Support Tool}\\
	We aim to release the DST for the planning of domiciliary care to make the software readily available to health care teams to assist with creating routes and schedules for staff. This will also allow us to obtain feedback from the users and a comparison with previous schedules.
	\item \textbf{Publications}\\
	We endeavour to submit two articles for publication: a methodological paper covering findings regarding exact methods and adaptations to stochastic demand, which will be submitted to a mathematical journal, and a scientific paper for a healthcare journal that will present our work on the efficiency of our DST in a real setting in comparison to business-as-usual planning. 
\end{enumerate}

%\begin{itemize}
%	\item Release DST
%	\item Second paper with exact method and stochastic demand
%	\item publish paper comparing abicare to a healthcare journal
%	\item Resubmit paper to computers and OR
%\end{itemize}

\section*{Timeline}
Below is a guide for the remainder of the project, and outlines how the above targets will be achieved:

\begin{enumerate}[leftmargin=*]
	\item \textbf{18th October 2021 -- 31st October 2021}\\
	Finalise initial publication and prepare to submit to journal.
	\item \textbf{November 2021}\\
	Release the DST. Prepare a questionnaire for nurses and planners in industry, which will be used as a basis for comparison after the DST has been used for a period of time. Meet with Abicare to discuss the frequency in which the DST should be used, and how results should be collected for future review. Complete formal documentation for the DST.
	\item \textbf{December 2021 -- January 2022}\\
	Focus on adapting the DST for stochastic demand. Identify the situations that should be considered, and attempt to model the changes.
	\item \textbf{January 2022 -- February 2022}\\
	Turn attention to exact methods. Study the literature to determine methods that will/not be suitable for our DST, and develop a model that can be adapted for the problem.
	\item \textbf{March 2022}\\
	Obtain results from Abicare. Send out final questionnaire to nurses and planners and compare with initial questionnaire. Compare and analyse outputs.
	\item \textbf{April 2022 -- May 2022}\\
	Gather data and test stochastic demand and exact method changes to the model.
	\item \textbf{June 2022}\\
	Potentially release new version of DST including stochastic demand and exact method. Prepare articles for submission to mathematical and healthcare journals. Conclude project.
\end{enumerate}

\end{document}