### 2021-12-18 tex/Documentation/Documentation.tex
updated documentation.

\documentclass[a4paper]{article}

\input{includes/preamble.tex}
\usepackage{parskip}

\begin{document}
\title{\vspace{-25mm}Quick Start Guide: Carrot DST\vspace{-15mm}}
%\author{Asyl L. Hawa}
\date{}
\maketitle

This is a quick start guide for the Carrot (CARe ROuting Tool) DST. It will allow the user to run the program for a given input file using default settings.

\section{Quick Start Guide}
\begin{enumerate}[leftmargin=*]
	\item Double-click on the \texttt{runprogram.bat} file to start the program. Two command prompts will appear:
	\begin{enumerate}
		\item The first window is for OSRM (Open Source Routing Machine). The last line in the window should be `[info] running and waiting for requests'. This line may take a few moments to appear.
		\item The second window will show some outputs from the program, and will initially appear blank.
	\end{enumerate}
	\item A third window will then appear titled `User Input Parameters'. It may take some time for this window to pop up due to the packaging of the program. An image of this window is provided below.
	\item In this window, simply select the desired area name in the corresponding field. If preferred, the time limit can also be altered -- currently it is set to 60 seconds. For `Select input file', use the `Browse' button to select the desired Excel input file containing data regarding clients and carers. The input file must be in the same format as the original Abicare file, of which a copy has been provided in the folder for reference. Finally, select a date for which a schedule and route is required using the `Calendar' button. All other fields may remain as they are.
	\item Double-click `OK' on the main `User Input Parameters' window, which will start the algorithm. Another window will appear outlining the progress of the program. Some outputs from the program will also be printed to the second command prompt window.
	\item When the program has terminated, all windows will close except the OSRM command prompt. This window can be closed manually on completion of the program.
	\item All outputs from the program will be stored in a new folder titled `outputs` in the main CarrotDST folder.
\end{enumerate}

\begin{figure}[H]
	\centering
	\includegraphics[width=0.72\linewidth]{figures/userinput}%
	\caption{User Input Parameters window.}
\end{figure}

\end{document}