### 2021-12-18 tex/Paper/combined_table_aith_results.tex
updated files.

\begin{table}[htbp!]
\centering
\caption{Performance of the best and average results found by GRASP DST compared with the three algorithms presented in \citet{aithaddadene2016} (ILS, \emph{ah\_ils}, basic GRASP,  \emph{ah\_grasp}, and GRASPxILS, \emph{ah\_gi}) for the three instance groups \emph{G1}, \emph{G2} and \emph{G3}. Figures in bold indicate the best results for each instance group.}
\begin{tabular}{lcrrrr}
	\toprule
	Result & Group & \emph{ah\_ils} & \emph{ah\_grasp} & \emph{ah\_gi} & \emph{GRASP} \\  \midrule                                            
    \emph{Best} & \emph{G1} & 38.16 & \textbf{38.04} & \textbf{38.04} & \textbf{38.04} \\  
    & \emph{G2} & -97.83 & -97.54 & -99.57 & \textbf{-99.69} \\  
    & \emph{G3} & -251.35 & -231.54 & -245.42 & \textbf{-251.55} \\  
    \midrule                                              
    \emph{Average} & \emph{G1} & 42.27 & 38.39 & 38.05 & \textbf{38.04} \\  
    & \emph{G2} & -91.85 & -94.04 & -97.91 & \textbf{-99.19} \\  
    & \emph{G3} & -240.82 & -226.79 & -239.44 & \textbf{-247.01} \\  
    \bottomrule
    \end{tabular}
\label{table:grouptableaithcombined}
\end{table}