### 2021-12-18 tex/Paper/figures/aithgantt18-4-l-2a.tex
new tikz figures.

\documentclass[usenames,dvipsnames,tikz]{standalone}
\usetikzlibrary{patterns}
%\usetikzlibrary{shapes.geometric}
%\usepackage{xcolor}
\colorlet{tBlue}{RoyalBlue!35!Cerulean}
\colorlet{tRed}{Red}
%\definecolor{tGreen}{HTML}{569909}
%\definecolor{tOrange}{HTML}{FA7602}
\usepackage{tikz}
\usepackage{standalone}
\begin{document}

\begin{tikzpicture}
%\draw [help lines] (-1,-1) grid (15,8);

% Nurse 0
\draw [thick] (0,6) rectangle (13.9, 7);
\draw [thick, pattern = north west lines, pattern color=black!30!white] (0,6) rectangle (0.22,7);
\draw [thick] (0.22,6) -- (0.22,7); % start of job 10
\draw [thick] (1.98,6) -- (1.98,7); % end of job 10
\draw [thick, pattern = north west lines, pattern color=black!30!white] (1.98,6) rectangle (2.14,7);
\draw [thick] (2.14,6) -- (2.14,7); % start of job 14
\draw [thick] (3.4,6) -- (3.4,7); % end of job 14
\draw [thick, pattern = north west lines, pattern color=black!30!white] (3.4,6) rectangle (3.68,7);
\draw [thick] (3.68,6) -- (3.68,7); % start of job 3
\draw [thick] (5.72,6) -- (5.72,7); % end of job 3
\draw [thick, pattern = north west lines, pattern color=black!30!white] (5.72,6) rectangle (5.86,7);
\draw [thick] (5.86,6) -- (5.86,7); % start of job 11
\draw [thick] (6.8,6) -- (6.8,7); % end of job 11
\draw [thick, pattern = north west lines, pattern color=black!30!white] (6.8,6) rectangle (7.32,7);
\draw [thick] (7.32,6) -- (7.32,7); % start of job 17
\draw [thick] (9.28,6) -- (9.28,7); % end of job 17
\draw [thick, pattern = north west lines, pattern color=black!30!white] (9.28,6) rectangle (9.52,7);
\draw [thick] (9.52,6) -- (9.52,7); % start of job 2
\draw [thick] (10.16,6) -- (10.16,7); % end of job 2
\draw [thick, pattern = north west lines, pattern color=black!30!white] (10.16,6) rectangle (10.22,7);
\draw [thick] (10.22,6) -- (10.22,7); % start of job 0 (DS)
\draw [thick] (13,6) -- (13,7); % end of job 0 (DS)
\draw [thick, pattern = north west lines, pattern color=black!30!white] (13,6) rectangle (13.42,7);
\draw [thick] (13.42,6) -- (13.42,7); % end of shift
\draw [thick, fill=black!20!white] (13.42,6) rectangle (13.9,7);

\node [right] at (-0.1,7.2) {\scriptsize{Nurse 0}};
\node at (1.1,6.6) {\scriptsize{job}}; 
\node at (1.1,6.35) {\scriptsize{10}};
\node at (2.77,6.6) {\scriptsize{job}};
\node at (2.77,6.35) {\scriptsize{14}};
\node at (4.7,6.6) {\scriptsize{job}};
\node at (4.7,6.35) {\scriptsize{3}};
\node at (6.33,6.6) {\scriptsize{job}};
\node at (6.33,6.35) {\scriptsize{11}};
\node at (8.3,6.6) {\scriptsize{job}};
\node at (8.3,6.35) {\scriptsize{17}};
\node at (9.84,6.6) {\scriptsize{job}};
\node at (9.84,6.35) {\scriptsize{2}};
\node at (11.61,6.6) {\scriptsize{job}};
\node at (11.61,6.35) {\scriptsize{0 (DS)}};

% Nurse 1
\draw [thick] (0,4) rectangle (13.9, 5);
\draw [thick, pattern = north west lines, pattern color=black!30!white] (0,4) rectangle (0.26,5);
\draw [thick] (0.26,4) -- (0.26,5); % start of job 16
\draw [thick] (1.64,4) -- (1.64,5); % end of job 16
\draw [thick, pattern = north west lines, pattern color=black!30!white] (1.64,4) rectangle (1.96,5);
\draw [thick] (1.96,4) -- (1.96,5); % start of job 15
\draw [thick] (3,4) -- (3,5); % end of job 15
\draw [thick, pattern = north west lines, pattern color=black!30!white] (3,4) rectangle (3.38,5);
\draw [thick] (3.38,4) -- (3.38,5); % start of job 1
\draw [thick] (4.12,4) -- (4.12,5); % end of job 1
\draw [thick, pattern = north west lines, pattern color=black!30!white] (4.12,4) rectangle (4.52,5);
\draw [thick] (4.52,4) -- (4.52,5); % start of job 7
\draw [thick] (6.66,4) -- (6.66,5); % end of job 7
\draw [thick, pattern = north west lines, pattern color=black!30!white] (6.66,4) rectangle (6.72,5);
\draw [thick] (6.72,4) -- (6.72,5); % start of job 12
\draw [thick] (9.08,4) -- (9.08,5); % end of job 12
\draw [thick, pattern = north west lines, pattern color=black!30!white] (9.08,4) rectangle (9.12,5);
\draw [thick] (9.12,4) -- (9.12,5); % ARRIVAL at job 18 (waiting time!)
\draw [thick, pattern = north west lines, pattern color=tRed] (9.12,4) rectangle (9.32,5);
\draw [thick] (9.32,4) -- (9.32,5); % start of job 18
\draw [thick] (9.96,4) -- (9.96,5); % end of job 18
\draw [thick, pattern = north west lines, pattern color=black!30!white] (9.96,4) rectangle (10.02,5);
\draw [thick] (10.02,4) -- (10.02,5); % ARRIVAL at job 0 (DS) (waiting time!)
\draw [thick, pattern = north west lines, pattern color=tRed] (10.02,4) rectangle (10.22,5);
\draw [thick] (10.22,4) -- (10.22,5); % start of job 0 (DS)
\draw [thick] (13,4) -- (13,5); % end of job 0 (DS)
\draw [thick, pattern = north west lines, pattern color=black!30!white] (13,4) rectangle (13.42,5);
\draw [thick] (13.42,4) -- (13.42,5); % end of shift
\draw [thick, fill=black!20!white] (13.42,4) rectangle (13.9,5);

\node [right] at (-0.1,5.2) {\scriptsize{Nurse 1}};
\node at (0.95,4.6) {\scriptsize{job}}; 
\node at (0.95,4.35) {\scriptsize{16}};
\node at (2.48,4.6) {\scriptsize{job}}; 
\node at (2.48,4.35) {\scriptsize{15}};
\node at (3.75,4.6) {\scriptsize{job}}; 
\node at (3.75,4.35) {\scriptsize{1}};
\node at (5.59,4.6) {\scriptsize{job}}; 
\node at (5.59,4.35) {\scriptsize{7}};
\node at (7.9,4.6) {\scriptsize{job}}; 
\node at (7.9,4.35) {\scriptsize{12}};
\node at (9.64,4.6) {\scriptsize{job}}; 
\node at (9.64,4.35) {\scriptsize{18}};
\node at (11.61,4.6) {\scriptsize{job}}; 
\node at (11.61,4.35) {\scriptsize{0 (DS)}};

% Nurse 2
\draw [thick] (0,2) rectangle (13.9, 3);
\draw [thick, pattern = north west lines, pattern color=black!30!white] (0,2) rectangle (0.26,3);
\draw [thick] (0.26,2) -- (0.26,3); % start of job 8
\draw [thick] (1,2) -- (1,3); % end of job 8
\draw [thick, pattern = north west lines, pattern color=black!30!white] (1,2) rectangle (1.18,3);
\draw [thick] (1.18,2) -- (1.18,3); % start of job 13
\draw [thick] (3.8,2) -- (3.8,3); % end of job 13
\draw [thick, pattern = north west lines, pattern color=black!30!white] (3.8,2) rectangle (3.92,3);
\draw [thick] (3.92,2) -- (3.92,3); % start of job 4
\draw [thick] (5.36,2) -- (5.36,3); % end of job 4
\draw [thick, pattern = north west lines, pattern color=black!30!white] (5.36,2) rectangle (5.66,3);
\draw [thick] (5.66,2) -- (5.66,3); % start of job 9
\draw [thick] (6.16,2) -- (6.16,3); % end of job 9
\draw [thick, pattern = north west lines, pattern color=black!30!white] (6.16,2) rectangle (6.34,3);
\draw [thick] (6.34,2) -- (6.34,3); % end of shift
\draw [thick,fill=black!20!white] (6.34,2) rectangle (13.9,3);

\node [right] at (-0.1,3.2) {\scriptsize{Nurse 2}};
\node at (0.63,2.6) {\scriptsize{job}}; 
\node at (0.63,2.35) {\scriptsize{8}};
\node at (2.49,2.6) {\scriptsize{job}}; 
\node at (2.49,2.35) {\scriptsize{13}};
\node at (4.64,2.6) {\scriptsize{job}}; 
\node at (4.64,2.35) {\scriptsize{4}};
\node at (5.91,2.6) {\scriptsize{job}}; 
\node at (5.91,2.35) {\scriptsize{9}};

% Nurse 3
\draw [thick] (0,0) rectangle (13.9, 1);
\draw [thick, pattern = north west lines, pattern color=black!30!white] (0,0) rectangle (0.36,1);
\draw [thick] (0.36,0) -- (0.36,1); % start of job 6
\draw [thick] (1.34,0) -- (1.34,1); % end of job 6
\draw [thick, pattern = north west lines, pattern color=black!30!white] (1.34,0) rectangle (1.58,1);
\draw [thick] (1.58,0) -- (1.58,1); % start of job 5
\draw [thick] (3.08,0) -- (3.08,1); % end of job 5
\draw [thick, pattern = north west lines, pattern color=black!30!white] (3.08,0) rectangle (3.22,1);
\draw [thick] (3.22,0) -- (3.22,1); % end of shift
\draw [thick, fill=black!20!white] (3.22,0) rectangle (13.9,1);

\node [right] at (-0.1,1.2) {\scriptsize{Nurse 3}};
\node at (0.85,0.6) {\scriptsize{job}}; 
\node at (0.85,0.35) {\scriptsize{6}};
\node at (2.33,0.6) {\scriptsize{job}}; 
\node at (2.33,0.35) {\scriptsize{5}};

\end{tikzpicture}

\end{document}

%\draw [thick] (0,6) rectangle (6.95, 7);
%\draw [thick] (0.11,6) -- (0.11,7); % start of job 10
%\draw [thick] (0.99,6) -- (0.99,7); % end of job 10
%\draw [thick] (1.07,6) -- (1.07,7); % start of job 14
%\draw [thick] (1.7,6) -- (1.7,7); % end of job 14
%\draw [thick] (1.84,6) -- (1.84,7); % start of job 3
%\draw [thick] (2.86,6) -- (2.86,7); % end of job 3
%\draw [thick] (2.93,6) -- (2.93,7); % start of job 11
%\draw [thick] (3.4,6) -- (3.4,7); % end of job 11
%\draw [thick] (3.66,6) -- (3.66,7); % start of job 17
%\draw [thick] (4.64,6) -- (4.64,7); % end of job 17
%\draw [thick] (4.76,6) -- (4.76,7); % start of job 2
%\draw [thick] (5.08,6) -- (5.08,7); % end of job 2
%\draw [thick] (5.11,6) -- (5.11,7); % start of job 0 (DS)
%\draw [thick] (6.5,6) -- (6.5,7); % end of job 0 (DS)
%\draw [thick] (6.71,6) -- (6.71,7); % end of shift