### 2021-12-18 tex/Paper/figures/aithgantt18-4-l-2c.tex

\documentclass[usenames,dvipsnames,tikz]{standalone}
\usetikzlibrary{patterns}
%\usetikzlibrary{shapes.geometric}
%\usepackage{xcolor}
\colorlet{tBlue}{RoyalBlue!35!Cerulean}
\colorlet{tRed}{Red}
%\definecolor{tGreen}{HTML}{569909}
%\definecolor{tOrange}{HTML}{FA7602}
\usepackage{tikz}
\usepackage{standalone}
\begin{document}

\begin{tikzpicture}
%\draw [help lines] (-1,-1) grid (15,8);

% Nurse 0
\draw [thick] (0,6) rectangle (14.04, 7);
\draw [thick, pattern = north west lines, pattern color=black!30!white] (0,6) rectangle (0.24,7);
\draw [thick] (0.24,6) -- (0.24,7); % start of job 12
\draw [thick] (2.10,6) -- (2.10,7); % end of job 12
\draw [thick, pattern = north west lines, pattern color=black!30!white] (2.10,6) rectangle (2.28,7);
\draw [thick] (2.28,6) -- (2.28,7); % start of job 16
\draw [thick] (3.76,6) -- (3.76,7); % end of job 16
\draw [thick, pattern = north west lines, pattern color=black!30!white] (3.76,6) rectangle (3.94,7);
\draw [thick] (3.94,6) -- (3.94,7); % start of job 14
\draw [thick] (5.50,6) -- (5.50,7); % end of job 14
\draw [thick, pattern = north west lines, pattern color=black!30!white] (5.50,6) rectangle (6.06,7);
\draw [thick] (6.06,6) -- (6.06,7); % ARRIVE at job 3
\draw [thick, pattern = north east lines, pattern color=tRed] (6.06,6) rectangle (6.96,7);
\draw [thick] (6.96,6) -- (6.96,7); % start of job 3
\draw [thick] (8.12,6) -- (8.12,7); % end of job 3
\draw [thick, pattern = north west lines, pattern color=black!30!white] (8.12,6) rectangle (8.48,7);
\draw [thick] (8.48,6) -- (8.48,7); % start of job 17
\draw [thick] (10.72,6) -- (10.72,7); % end of job 17
\draw [thick, pattern = north west lines, pattern color=black!30!white] (10.72,6) rectangle (10.98,7);
\draw [thick] (10.98,6) -- (10.98,7); % start of job 1
\draw [thick] (12.94,6) -- (12.94,7); % end of job 1
\draw [thick, pattern = north west lines, pattern color=black!30!white] (12.94,6) rectangle (13.18,7);
\draw [thick] (13.18,6) -- (13.18,7); % end of shift
\draw [thick, fill=black!20!white] (13.18,6) rectangle (14.04,7);

\node [right] at (-0.1,7.2) {\scriptsize{Nurse 0}};
\node at (1.17,6.6) {\scriptsize{job}}; 
\node at (1.17,6.35) {\scriptsize{12}};
\node at (3.02,6.6) {\scriptsize{job}};
\node at (3.02,6.35) {\scriptsize{16}};
\node at (4.72,6.6) {\scriptsize{job}};
\node at (4.72,6.35) {\scriptsize{14}};
\node at (7.54,6.6) {\scriptsize{job}};
\node at (7.54,6.35) {\scriptsize{3(DS)}};
\node at (9.60,6.6) {\scriptsize{job}};
\node at (9.60,6.35) {\scriptsize{17}};
\node at (11.96,6.6) {\scriptsize{job}};
\node at (11.96,6.35) {\scriptsize{1}};

% Nurse 1
\draw [thick] (0,4) rectangle (14.04, 5);
\draw [thick, pattern = north west lines, pattern color=black!30!white] (0,4) rectangle (0.34,5);
\draw [thick] (0.34,4) -- (0.34,5); % ARRIVE at job 9
\draw [thick, pattern = north east lines, pattern color=tRed] (0.34,4) rectangle (4.18,5);
\draw [thick] (4.18,4) -- (4.18,5); % start of job 9
\draw [thick] (5.48,4) -- (5.48,5); % end of job 9
\draw [thick, pattern = north west lines, pattern color=black!30!white] (5.48,4) rectangle (6.10,5);
\draw [thick] (6.10,4) -- (6.10,5); % start of job 8
\draw [thick] (7.86,4) -- (7.86,5); % end of job 8
\draw [thick, pattern = north west lines, pattern color=black!30!white] (7.86,4) rectangle (8.27,5);
\draw [thick] (8.27,4) -- (8.27,5); % end of shift
\draw [thick, fill=black!20!white] (8.27,4) rectangle (14.04,5);

\node [right] at (-0.1,5.2) {\scriptsize{Nurse 1}};
\node at (4.83,4.6) {\scriptsize{job}}; 
\node at (4.83,4.35) {\scriptsize{9(DS)}};
\node at (6.98,4.6) {\scriptsize{job}}; 
\node at (6.98,4.35) {\scriptsize{8}};

% Nurse 2
\draw [thick] (0,2) rectangle (14.04, 3);
\draw [thick, pattern = north west lines, pattern color=black!30!white] (0,2) rectangle (0.58,3);
\draw [thick] (0.58,2) -- (0.58,3); % start of job 4
\draw [thick] (2.44,2) -- (2.44,3); % end of job 4
\draw [thick, pattern = north west lines, pattern color=black!30!white] (2.44,2) rectangle (2.88,3);
\draw [thick] (2.88,2) -- (2.88,3); % start of job 15
\draw [thick] (4.10,2) -- (4.10,3); % end of job 15
\draw [thick, pattern = north west lines, pattern color=black!30!white] (4.10,2) rectangle (4.18,3);
\draw [thick] (4.18,2) -- (4.18,3); % start of job 9
\draw [thick] (5.48,2) -- (5.48,3); % end of job 9
\draw [thick, pattern = north west lines, pattern color=black!30!white] (5.48,2) rectangle (5.86,3);
\draw [thick] (5.86,2) -- (5.86,3); % start of job 13
\draw [thick] (6.84,2) -- (6.84,3); % end of job 13
\draw [thick, pattern = north west lines, pattern color=black!30!white] (6.84,2) rectangle (6.96,3);
\draw [thick] (6.96,2) -- (6.96,3); % start of job 3
\draw [thick] (8.12,2) -- (8.12,3); % end of job 3
\draw [thick, pattern = north west lines, pattern color=black!30!white] (8.12,2) rectangle (8.52,3);
\draw [thick] (8.52,2) -- (8.52,3); % end of shift
\draw [thick,fill=black!20!white] (8.52,2) rectangle (14.04,3);

\node [right] at (-0.1,3.2) {\scriptsize{Nurse 2}};
\node at (1.51,2.6) {\scriptsize{job}}; 
\node at (1.51,2.35) {\scriptsize{4}};
\node at (3.49,2.6) {\scriptsize{job}}; 
\node at (3.49,2.35) {\scriptsize{15}};
\node at (4.83,2.6) {\scriptsize{job}}; 
\node at (4.83,2.35) {\scriptsize{9(DS)}};
\node at (6.35,2.6) {\scriptsize{job}}; 
\node at (6.35,2.35) {\scriptsize{13}};
\node at (7.54,2.6) {\scriptsize{job}}; 
\node at (7.54,2.35) {\scriptsize{3(DS)}};

% Nurse 3
\draw [thick] (0,0) rectangle (14.04, 1);
\draw [thick, pattern = north west lines, pattern color=black!30!white] (0,0) rectangle (0.18,1);
\draw [thick] (0.18,0) -- (0.18,1); % start of job 6
\draw [thick] (1.34,0) -- (1.34,1); % end of job 6
\draw [thick, pattern = north west lines, pattern color=black!30!white] (1.34,0) rectangle (1.50,1);
\draw [thick] (1.50,0) -- (1.50,1); % start of job 5
\draw [thick] (3.52,0) -- (3.52,1); % end of job 5
\draw [thick, pattern = north west lines, pattern color=black!30!white] (3.52,0) rectangle (4.04,1);
\draw [thick] (4.04,0) -- (4.04,1); % start of job 5
\draw [thick] (5.20,0) -- (5.20,1); % end of job 5
\draw [thick, pattern = north west lines, pattern color=black!30!white] (5.20,0) rectangle (5.26,1);
\draw [thick] (5.26,0) -- (5.26,1); % start of job 5
\draw [thick] (7.68,0) -- (7.68,1); % end of job 5
\draw [thick, pattern = north west lines, pattern color=black!30!white] (7.68,0) rectangle (7.88,1);
\draw [thick] (7.88,0) -- (7.88,1); % start of job 5
\draw [thick] (9.26,0) -- (9.26,1); % end of job 5
\draw [thick, pattern = north west lines, pattern color=black!30!white] (9.26,0) rectangle (9.60,1);
\draw [thick] (9.60,0) -- (9.60,1); % start of job 5
\draw [thick] (11.58,0) -- (11.58,1); % end of job 5
\draw [thick, pattern = north west lines, pattern color=black!30!white] (11.58,0) rectangle (11.60,1);
\draw [thick] (11.60,0) -- (11.60,1); % start of job 5
\draw [thick] (13.48,0) -- (13.48,1); % end of job 5
\draw [thick, pattern = north west lines, pattern color=black!30!white] (13.48,0) rectangle (13.78,1);
\draw [thick] (13.78,0) -- (13.78,1); % end of shift
\draw [thick, fill=black!20!white] (13.78,0) rectangle (14.04,1);

\node [right] at (-0.1,1.2) {\scriptsize{Nurse 3}};
\node at (0.76,0.6) {\scriptsize{job}}; 
\node at (0.76,0.35) {\scriptsize{6}};
\node at (2.51,0.6) {\scriptsize{job}}; 
\node at (2.51,0.35) {\scriptsize{5}};
\node at (4.62,0.6) {\scriptsize{job}}; 
\node at (4.62,0.35) {\scriptsize{0}};
\node at (6.47,0.6) {\scriptsize{job}}; 
\node at (6.47,0.35) {\scriptsize{7}};
\node at (8.57,0.6) {\scriptsize{job}}; 
\node at (8.57,0.35) {\scriptsize{11}};
\node at (10.59,0.6) {\scriptsize{job}}; 
\node at (10.59,0.35) {\scriptsize{2}};
\node at (12.54,0.6) {\scriptsize{job}}; 
\node at (12.54,0.35) {\scriptsize{10}};

\end{tikzpicture}

\end{document}

%\draw [thick] (0,6) rectangle (6.95, 7);
%\draw [thick] (0.11,6) -- (0.11,7); % start of job 10
%\draw [thick] (0.99,6) -- (0.99,7); % end of job 10
%\draw [thick] (1.07,6) -- (1.07,7); % start of job 14
%\draw [thick] (1.7,6) -- (1.7,7); % end of job 14
%\draw [thick] (1.84,6) -- (1.84,7); % start of job 3
%\draw [thick] (2.86,6) -- (2.86,7); % end of job 3
%\draw [thick] (2.93,6) -- (2.93,7); % start of job 11
%\draw [thick] (3.4,6) -- (3.4,7); % end of job 11
%\draw [thick] (3.66,6) -- (3.66,7); % start of job 17
%\draw [thick] (4.64,6) -- (4.64,7); % end of job 17
%\draw [thick] (4.76,6) -- (4.76,7); % start of job 2
%\draw [thick] (5.08,6) -- (5.08,7); % end of job 2
%\draw [thick] (5.11,6) -- (5.11,7); % start of job 0 (DS)
%\draw [thick] (6.5,6) -- (6.5,7); % end of job 0 (DS)
%\draw [thick] (6.71,6) -- (6.71,7); % end of shift